Methods for Large Scale Learning to Rank

### 2014-11-04 report/Chapters/CrossBenchmarkComparison.tex
kleine stabiliteitstweak clustercommunicatie + betere uitleg ListNet

LARF and LRUF score very high \ac{NWN} scores, but based on only relatively few data sets, just as in the \ac{NDCG}-comparisons. Notable is that SmoothRank and ListNet, which showed both high accuracy and high certainty on all \ac{NDCG}-comparisons, are not within the best performing methods in the \ac{MAP}-comparison. A deeper look in the raw data Tables \ref{tab:raw_data_norm_winnum_ndcg35} and \ref{tab:raw_data_norm_winnum_ndcg10map} in Appendices \ref{app:norm_winnum_ndcg35} and \ref{app:norm_winnum_ndcg10map} respectively shows that LAC-MR-OR is evaluated on many more data sets for \ac{MAP} compared to \ac{NDCG}, which resulted in LAC-MR-OR obtaining equal certainty to ListNet with a higher \ac{NWN}. SmoothRank performed a \ac{NWN} of around 0.53 over 7 data sets, which is still good in both certainty and accuracy, but not among the top methods. RE-QR is one of the best performers in the \ac{MAP}-comparison with a reasonable amount of benchmark evaluations. No reported \ac{NDCG} performance was found in the literature study for RE-QR. There is a lot of certainty on the accuracy of RankBoost and Rank\ac{SVM} as both models are evaluated on the majority of data sets included in the comparison for the \ac{MAP}-metric, but given their \ac{NWN} it can said that both methods are not within the top performing Learning to Rank methods.

>>>>

LARF and LRUF score very high \ac{NWN} scores, but based on only relatively few data sets, just as in the \ac{NDCG}-comparisons. Notable is that SmoothRank and ListNet, which showed both high accuracy and high certainty on all \ac{NDCG}-comparisons, are not within the best performing methods in the \ac{MAP}-comparison. A deeper look in the raw data Tables \ref{tab:raw_data_norm_winnum_ndcg35} and \ref{tab:raw_data_norm_winnum_ndcg10map} in Appendices \ref{app:norm_winnum_ndcg35} and \ref{app:norm_winnum_ndcg10map} respectively shows that LAC-MR-OR is evaluated on many more data sets for \ac{MAP} compared to \ac{NDCG}, which resulted in LAC-MR-OR obtaining equal certainty to ListNet with a higher \ac{NWN}. SmoothRank performed a \ac{NWN} of around 0.53 over 7 data sets, which is still good in both certainty and accuracy, but not among the top methods. RE-QR is one of the best performers in the \ac{MAP}-comparison with a reasonable amount of benchmark evaluations. No reported \ac{NDCG} performance was found in the literature study for RE-QR. There is a lot of certainty on the accuracy of RankBoost and Rank\acs{SVM} as both models are evaluated on the majority of data sets included in the comparison for the \ac{MAP}-metric, but given their \ac{NWN} it can said that both methods are not within the top performing Learning to Rank methods.

### 2014-11-04 report/Chapters/LTRMethods.tex
kleine stabiliteitstweak clustercommunicatie + betere uitleg ListNet

\section{ListNet}
\label{sec:ltrmethods_listnet}
ListNet \cite{Cao2007} is a listwise ranking function whose loss function is not directly related to information retrieval evaluation metrics. ListNet's loss function is defined using a probability distribution on permutations. Probability distributions on permutations have been a research topic within the field of probability theory which has been extensively researched. ListNet uses the Plackett-Luce model \cite{Plackett1975,Luce1959}, which is one of the most well-known permutation probability distributions. The Plackett-Luce model defines the probability of all possible permutations $\pi$, given all document ranking scores $S$, as shown in Equation \ref{eq:plackett_luce}.
\begin{equation}
P(\pi|S) = \prod\limits_{j=1}^{m}\frac{\phi(s_{\pi^{-1}(j)})}{\sum\nolimits_{u=1}^{m}\phi(s_{\pi^{-1}(u)})}
\label{eq:plackett_luce}
\end{equation}
ListNet uses Gradient Descent to optimise a neural network such that its Cross Entropy loss compared to the Plackett-Luce distribution over the training data relevance labels is minimal. Note that some sources, including Liu \cite{Liu2007}, describe ListNet as using \ac{KL divergence} as loss function. \ac{KL divergence} and Cross Entropy are however identical up to an additive constant when comparing distribution $q$ against a fixed reference distribution $p$. This follows from the definition of Cross Entropy Loss, which is shown in Equation \ref{eq:cross_entropy}.
\begin{equation}
H(p,q) = H(p) + D_{KL}(p||q)
\label{eq:cross_entropy}
\end{equation}
where $H(p)$ is the entropy of $p$ and $D_{KL}(p||q)$ is the \ac{KL divergence} of $q$ from $p$.\\

\noindent Equation \ref{eq:gradient_descent} describes the gradient descent step to minimise loss function $L(y^{(i)},z^{(i)}(f_\omega))$ with respect to parameter $\omega$. 
\begin{equation}
\Delta\omega = \frac{\partial L(y^{(i)},z^{(i)}(f_\omega))}{\partial \omega} = - \sum_{\forall g \in \mathscr{G}_k}\limits\frac{\partial P_{z^{(i)}(f_\omega)}(g)}{\partial \omega}\frac{P_{y^{(i)}}(g)}{P_{z^{(i)}(f_\omega)}(g)}
\label{eq:gradient_descent}
\end{equation}

>>>>

\section{ListNet}
\label{sec:ltrmethods_listnet}
ListNet \cite{Cao2007} is a listwise ranking function whose loss function is not directly related to information retrieval evaluation metrics. ListNet's loss function is defined using a probability distribution on permutations. Probability distributions on permutations have been a research topic within the field of probability theory which has been extensively researched. ListNet is based on the Plackett-Luce model \cite{Plackett1975,Luce1959}, which is a permutation probability distribution over permutations that is often used in the field of econometrics. The Plackett-Luce model defines a probability over permutations $\pi$ from the set of all possible permutations $\Omega$, given all document ranking scores $s$. The Plackett-Luce model defines the probability of a permutation $\pi$ given the list as scores $s$ as shown in Equation \ref{eq:plackett_luce}.
\begin{equation}
P_s(\pi) = \prod\limits_{j=1}^{n}\frac{\phi(s_{\pi(j)})}{\sum\nolimits_{k=j}^{n}\phi(s_{\pi(k)})}
\label{eq:plackett_luce}
\end{equation}
where $\pi$ is a permutation on the $n$ objects, $\phi$ is an monotonically increasing and strictly positive function.\\

However, the total number of permutations of a list of $n$ documents is $n!$, therefore the calculations of the probabilities of all possible permutations quickly becomes intractable. To cope with this problem, the authors of ListNet propose the calculation of a \emph{top one probability} as alternative. The top one probability of an object $j$ equals the sum of the permutations probabilities of permutations in which object $j$ is ranked as the top object. It seems to be the case that the $n!$ permutation probabilities still need to be calculated for this, but we can calculate the top one probability efficiently using Equation \ref{eq:top_one}.
\begin{equation}
P_s(j) = \frac{\phi{s_j}}{\sum\nolimits_{k=1}^{n}\phi(s_k)}
\label{eq:top_one}
\end{equation}

ListNet uses Gradient Descent to optimise a neural network such that its Cross Entropy loss compared to the top one over the training data relevance labels is minimal. For the monotonically increasing and strictly positive function $\phi$ is chosen to be the exponential function. Note that some sources, including Liu \cite{Liu2007}, describe ListNet as using \ac{KL divergence} as loss function. \ac{KL divergence} and Cross Entropy are however identical up to an additive constant when comparing distribution $q$ against a fixed reference distribution $p$. This follows from the definition of Cross Entropy Loss, which is shown in Equation \ref{eq:cross_entropy}.
\begin{equation}
L(y^{(i)},z^{(i)}) = \sum\limits_{j=1}^{n}P_{y^{(i)}}(j)\log(P_{z^{(i)}}(j))
\label{eq:cross_entropy}
\end{equation}
where $H(p)$ is the entropy of $p$ and $D_{KL}(p||q)$ is the \ac{KL divergence} of $q$ from $p$.\\

\noindent Equation \ref{eq:gradient_descent} describes the gradient descent step to minimise listwise loss function $L(y^{(i)},z^{(i)}(f_\omega))$ with respect to parameter $\omega$. 
\begin{multline}
\Delta\omega = \frac{\partial L(y^{(i)},z^{(i)}(f_\omega))}{\partial \omega} = - \sum\limits_{j=1}^{n^{(i)}}P_{y^{(i)}}(x_{j}^{(i)})\frac{\partial f_{\omega}(x_{j}^{(i)})}{\partial\omega} \\ 
+ \frac{1}{\sum\nolimits_{j=1}^{n^{(i)}}exp(f_{\omega}(x_{j}^{(i)}))} \sum\limits_{j=1}^{n^{(i)}}exp(f_{\omega}(x_{j}^{(i)})) \frac{\partial f_{\omega(x_{j}^{(i)})}}{\partial\omega}
\label{eq:gradient_descent}
\end{multline}

### 2014-11-04 report/Chapters/RelatedWork.tex
kleine stabiliteitstweak clustercommunicatie + betere uitleg ListNet

\subsection{Stacked generalisation}
% TODO: Kijken naar deze subsection, is erg zwak %
A \ac{DSN} is a processing architecture developed from the field of \emph{Deep Learning}, that uses the \ac{MSE} loss function that is easily optimisable. \ac{DSN} is based on the stacked generalization (stacking) ensemble learning method \cite{Wolpert1992}, an approach where several learning models are stacked on top of each other in such a way that the output of a model is used as one of the input features of models higher up in the stack. Each layer in the stack learns has the task of learning the weights of the inputs of that layer. The close-form constraints between input and output weights allow the input weight matrices to be estimated in a parallel manner. Deng et al. \cite{Deng2013} % TODO: beschrijven Deng et al.%\\

>>>>

\subsection{Stacked generalisation}
A \ac{DSN} is a processing architecture developed from the field of \emph{Deep Learning}. \ac{DSN}s are based on the stacked generalization (also known as stacking) ensemble learning method \cite{Wolpert1992}, which is a multi-layer approach where multiple learning models are stacked on top of each other in such a way that the outputs of the models are used as the input features for the models in the next layer of the stack. Stacked generalisation models are easy to compute in parallel, as the models within the same layer can be trained independently. Deng et al. \cite{Deng2013} used a deep architecture of stacked generalisation (\ac{DSN}) to make pointwise relevance label predictions such that the \ac{MSE} of the predicted relevance labels as compared to the ground truth relevance labels are optimised. No speed-up results of parallel computation are reported, although it is stated repeatedly that the method is scalable and easy to compute in parallel
